---
title: "Simulation Results"
author: "AMM"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/annettemolinaro/Repository/Aim2/Sims")

library(vioplot)

compdc.error<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_CompositeDC_ErrorToTruth.txt")
compthirds.error<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_CompositeThirds_ErrorToTruth.txt")

compRob.CVerror<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_CompositeRob_ErrorToTruth.txt")
compRob.errorU<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_CompositeRob_ErrorToTruth.errorU.lambdas.txt")
compRob.Error<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_CompositeRob_ErrorToTruth.ErrorOptimisim.Lambdas.txt")
compRob.error<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_CompositeRob_ErrorToTruth.error.lambdas.txt")

compdc.error.500<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_500_CompositeDC_ErrorToTruth.txt")
compthirds.error.500<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_500_CompositeThirds_ErrorToTruth.txt")
compRob.CVerror.500<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_500_CompositeRob_ErrorToTruth.txt")
compRob.errorU.500<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_500_CompositeRob_ErrorToTruth.errorU.lambdas.txt")
compRob.Error.500<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_500_CompositeRob_ErrorToTruth.ErrorOptimisim.Lambdas.txt")
compRob.error.500<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/OlderFiles/friedman1_500_CompositeRob_ErrorToTruth.error.lambdas.txt")

compdc.error.500.uplim<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeDC_Train500_uplim20_ErrorToTruth.txt")
compthirds.error.500.uplim<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeThirds_500_uplim_ErrorToTruth.txt")
compRob.CVerror.500.uplim<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeRob_500_uplim_ErrorToTruth.txt")
compRob.errorU.500.uplim<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeRob_500_uplim_ErrorToTruth.errorU.lambdas.txt")
compRob.Error.500.uplim<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeRob_500_uplim_ErrorToTruth.ErrorOptimisim.Lambdas.txt")
compRob.error.500.uplim<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeRob_500_uplim_ErrorToTruth.error.lambdas.txt")


compdc.error.500.uplim.lambda<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeDC_Train500_uplim20_MinErrorLambda.txt")[1:500,]
compthirds.error.500.uplim.lambda<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeThirds_500_uplim_MinErrorLambda.txt")
compRob.CVerror.500.uplim.lambda<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeRob_500_uplim_MinErrorLambda.txt")
compRob.errorU.500.uplim.lambda<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeRob_500_uplim_MinErrorLambda.errorU.lambdas.txt")
compRob.Error.500.uplim.lambda<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeRob_500_uplim_MinErrorLambda.ErrorOptimisim.Lambdas.txt")
compRob.error.500.uplim.lambda<-read.table("/Users/annettemolinaro/Repository/Aim2/Sims/friedman1_500_CompositeRob_500_uplim_MinErrorLambda.error.lambdas.txt")

```
## Composite Methods
* DC - Simple method we discussed in DC
* Thirds -- approach splitting data into 3
* R_CV -- uses CV error
* R_OpErr -- uses Op-Corr apparent error
* R_AppRes -- 'apparent error' using actual response
* R_AppCI -- 'apparent error' using ci response

## Friedman 1 - Simulation 

* $$X_{1}, \cdots,X_{9} \sim U(0,1)$$
* $$ Y= 10 \sin(\pi X_1\ast X_2)+20(X_{3}-.5)^2+10X_4+5X_{5}+\epsilon$$
* $$ \epsilon \sim N(0,1)$$


There are 500 simulations performed. 



## First scenario
### training set of size 250 and a test set of size 1000 
### Prediction error of difference between True outcome and predicted outcome.
```{r,echo=FALSE}

    xy<-cbind(compdc.error,compthirds.error,compRob.CVerror,compRob.Error,compRob.errorU,compRob.error)
    name.vec <- c("DC","Thirds","R_CV","R_OpErr","R_AppRes","R_AppCI")
    boxplot(xy,names=name.vec,ylab="Average Squared Error",main="Friedman 1")
    vioplot(as.vector(unlist(compdc.error)),as.vector(unlist(compthirds.error)),as.vector(unlist(compRob.CVerror)),as.vector(unlist(compRob.Error)),as.vector(unlist(compRob.errorU)),as.vector(unlist(compRob.error)),names=name.vec)
    
    tabl <- cbind(summary(compdc.error),summary(compthirds.error),summary(compRob.CVerror),summary(compRob.Error),summary(compRob.errorU),summary(compRob.error))
    colnames(tabl) <- name.vec
    #write.table(tabl,"Friedman1_table.txt",sep="\t",row.names=TRUE,col.names=TRUE,quote=FALSE)
    tabl
```


## Second scenario
### With 500 in the training set and 1000 in the test set
### Prediction error of difference between True outcome and predicted outcome.
```{r,echo=FALSE}

    xy.500<-cbind(unlist(compdc.error.500)[1:966],unlist(compthirds.error.500)[1:966],unlist(compRob.CVerror.500),unlist(compRob.Error.500),unlist(compRob.errorU.500),unlist(compRob.error.500))
    name.vec <- c("DC","Thirds","R_CV","R_OpErr","R_AppRes","R_AppCI")
    boxplot(xy.500,names=name.vec,ylab="Average Squared Error",main="Friedman 1")
    vioplot(as.vector(unlist(compdc.error.500)),as.vector(unlist(compthirds.error.500)),as.vector(unlist(compRob.CVerror.500)),as.vector(unlist(compRob.Error.500)),as.vector(unlist(compRob.errorU.500)),as.vector(unlist(compRob.error.500)),names=name.vec)
    
    tabl <- cbind(summary(compdc.error.500),summary(compthirds.error.500),summary(compRob.CVerror.500),summary(compRob.Error.500),summary(compRob.errorU.500),summary(compRob.error.500))
    colnames(tabl) <- name.vec
    #write.table(tabl,"Friedman1_table.txt",sep="\t",row.names=TRUE,col.names=TRUE,quote=FALSE)
    tabl
```



## Third scenario
### With a grid for lambda and 500 in the training set and 1000 in the test set  
### Prediction error of difference between True outcome and predicted outcome.
```{r,echo=FALSE}

    xy.500.uplim<-cbind(unlist(compdc.error.500.uplim)[1:73],unlist(compthirds.error.500.uplim)[1:73],unlist(compRob.CVerror.500.uplim),unlist(compRob.Error.500.uplim),unlist(compRob.errorU.500.uplim),unlist(compRob.error.500.uplim))
    name.vec <- c("DC","Thirds","R_CV","R_OpErr","R_AppRes","R_AppCI")
    boxplot(xy.500.uplim,names=name.vec,ylab="Average Squared Error",main="Friedman 1")
    vioplot(as.vector(unlist(compdc.error.500.uplim)[1:73]),as.vector(unlist(compthirds.error.500.uplim)[1:73]),as.vector(unlist(compRob.CVerror.500.uplim)),as.vector(unlist(compRob.Error.500.uplim)),as.vector(unlist(compRob.errorU.500.uplim)),as.vector(unlist(compRob.error.500.uplim)),names=name.vec)
    
    tabl <- cbind(summary(compdc.error.500.uplim),summary(compthirds.error.500.uplim),summary(compRob.CVerror.500.uplim),summary(compRob.Error.500.uplim),summary(compRob.errorU.500.uplim),summary(compRob.error.500.uplim))
    colnames(tabl) <- name.vec
    #write.table(tabl,"Friedman1_table.txt",sep="\t",row.names=TRUE,col.names=TRUE,quote=FALSE)
    tabl
```
 
## Third Scenario - now looking at lambda values
### With a grid for lambda and 500 in the training set and 1000 in the test set  
### Lambda chosen for each method (all limited to a grid of 0 to 20 for lambda with 100 points)
```{r, echo=FALSE}
      lambda.500.uplim<-cbind(unlist(compdc.error.500.uplim.lambda)[1:73],unlist(compthirds.error.500.uplim.lambda)[1:73],unlist(compRob.CVerror.500.uplim.lambda),unlist(compRob.Error.500.uplim.lambda),unlist(compRob.errorU.500.uplim.lambda),unlist(compRob.error.500.uplim.lambda))
    name.vec <- c("DC","Thirds","R_CV","R_OpErr","R_AppRes","R_AppCI")
    boxplot(lambda.500.uplim,names=name.vec,ylab="Average Squared Error",main="Friedman 1")
    
    
    vioplot(as.vector(unlist(compdc.error.500.uplim.lambda)[1:73]),as.vector(unlist(compthirds.error.500.uplim.lambda)[1:73]),as.vector(unlist(compRob.CVerror.500.uplim.lambda)),as.vector(unlist(compRob.Error.500.uplim.lambda)),as.vector(unlist(compRob.errorU.500.uplim.lambda)),as.vector(unlist(compRob.error.500.uplim.lambda)),names=name.vec)
    
    tabl <- cbind(summary(compdc.error.500.uplim.lambda),summary(compthirds.error.500.uplim.lambda),summary(compRob.CVerror.500.uplim.lambda),summary(compRob.Error.500.uplim.lambda),summary(compRob.errorU.500.uplim.lambda),summary(compRob.error.500.uplim.lambda))
    colnames(tabl) <- name.vec
    #write.table(tabl,"Friedman1_table.txt",sep="\t",row.names=TRUE,col.names=TRUE,quote=FALSE)
    tabl
```